### 2018-09-05 augsburg/talks/trento18/euroscipy.tex

\setbeamerfont{frametitle}{family=\sffamily\firamedium}
\setbeamercolor{alerted text}{fg=red!70!black}
\setbeamercolor{structure}{fg=Navy}

@@

@@

\begin{frame}{About me}
 \begin{itemize}
  \item theoretical physicist at the Universität Augsburg
  \item teaching programming to physicists and materials scientists since 2010
  \item involved in two Erasmus+ projects on education and computing
	\begin{itemize}
         \item iCSE4school (2015--2017)
	 \item Juypter@edu (2017--2019)
	\end{itemize}
 \end{itemize}

 \vspace{0.8truecm}
 \structure{Jupyter@edu}\qquad \raisebox{-0.2truecm}{\includegraphics[width=0.25\textwidth]{EU_flag-Erasmus__vect_POS}}

 \begin{itemize}
  \item University of Silesia, Poland (Leader)
  \item Universität Augsburg, Germany
  \item Universidade Portucalense Infante D. Henrique, Portugal
  \item European University Cyprus, Cyprus
 \end{itemize}
\end{frame}

\begin{frame}{Grading}
 \includegraphics[width=\textwidth]{function}

 \includegraphics[width=\textwidth]{tests}
\end{frame}

\begin{frame}{Functions}
 the students' code needs to be structured in functions to be tested
 \begin{itemize}
  \item concept of a function needs to be introduced very early in the course\\
	but we will always have well defined inputs and outputs\\
	the more special aspects of a function (no argument, default arguments,
	keyword arguments, \dots) are not required
  \item students get used to logically structured code from the very beginning\\
	but they do not go through the routine of structuring the code by themselves
  \item docstrings are required to define what the function is supposed to do\\
	students get used early to the idea that a function contains a docstring\\
	but they are not writing the docstring by themselves
 \end{itemize}
\end{frame}

\begin{frame}{Grading with tests}
 \begin{itemize}
  \item Test-driven development\\
	but: tests are not developed by the students
  \item Tests allow to give feedback through error messages\\
	but: students tend to rely on this feedback instead of developing their own critical view on their code\\
	potential problem of incomplete tests
  \item ``Stupid mistakes'' will be made even when students get more advanced\\
	also test for the simple things: are results returned, do they have the correct type, \dots\\
	``trivial'' standard tests + tests of specific functionality
  \item The test should not disclose the solution.
 \end{itemize}
\end{frame}

\begin{frame}{Coping with heterogeneity}
 \begin{itemize}
  \item students without programming experience and (sometimes) code writing apprehension
  \item students with knowledge in another programming language\\
	code often does not look very pythonic, example: loop over objects vs. loop over indices
 \end{itemize}

@@

 \begin{itemize}
  \item scientific applications are often considered as an extra mental burden
 \end{itemize}
\end{frame}

\begin{frame}{Examples of exercises}

@@

\begin{frame}{Branching paths}
 \only<1>{%
  \begin{columns}
   \begin{column}{0.3\textwidth}
    \includegraphics[height=0.8\textheight]{branching_1}
   \end{column}%
   \begin{column}{0.7\textwidth}
    \includegraphics[width=\textwidth]{JuliaSet_screenshot}
   \end{column}
  \end{columns}
 }%
 \only<2>{%
  \begin{columns}
   \begin{column}{0.3\textwidth}
    \includegraphics[height=0.8\textheight]{branching_2}
   \end{column}%
   \begin{column}{0.7\textwidth}
    \includegraphics[width=\textwidth]{ColorRepresentation_screenshot}
   \end{column}
  \end{columns}
 }%
 \only<3>{%
  \begin{columns}
   \begin{column}{0.3\textwidth}
    \includegraphics[height=0.8\textheight]{branching_3}
   \end{column}%
   \begin{column}{0.7\textwidth}
    \includegraphics[width=\textwidth]{Grid_screenshot}
   \end{column}
  \end{columns}
 }%
 \only<4>{%
  \begin{columns}
   \begin{column}{0.3\textwidth}
    \includegraphics[height=0.8\textheight]{branching_4}
   \end{column}%
   \begin{column}{0.7\textwidth}
    \includegraphics[width=\textwidth]{JuliaIteration_screenshot}
   \end{column}
  \end{columns}
 }
\end{frame}

\begin{frame}{Branching paths}
 \begin{itemize}
  \item individual path through problem possible
  \item other notebooks are opened in new tabs, it is easy to lose track
 \end{itemize}
\end{frame}

>>>>

\setbeamerfont{frametitle}{family=\sffamily\firamedium}
\setbeamercolor{alerted text}{fg=red!70!black}
\setbeamercolor{structure}{fg=Navy}
\setbeamertemplate{section in toc shaded}[default][30]

@@

\newcommand\but{\alert{\textit{but}} }

@@

\begin{frame}{About me}
 \begin{itemize}
  \item theoretical physicist at the Universität Augsburg
  \item teaching programming to physicists and materials scientists\\
	since 2010, typically 10+ students
  \item in two years: mandatory course for 100+ students $\rightarrow$
	  \structure{nbgrader}
  \item involved in two Erasmus+ projects on education and computing
	\begin{itemize}
         \item iCSE4school (2015--2017)
	 \item Juypter@edu (2017--2019)
	\end{itemize}
 \end{itemize}

 \vspace{0.8truecm}
 \structure{Jupyter@edu}\hspace{4.7truecm}%
 \raisebox{-0.7truecm}{\includegraphics[width=0.37\textwidth]{EU_flag-Erasmus__vect_POS}}

 \vspace{-0.7truecm}
 \begin{itemize}
  \item University of Silesia, Poland (Leader)
  \item Universität Augsburg, Germany
  \item Universidade Portucalense Infante D. Henrique, Portugal
  \item European University Cyprus, Cyprus
 \end{itemize}
\end{frame}

\begin{frame}{nbgrader}
 \begin{center}
 \includegraphics[width=\textwidth]{nbgrader}
 \end{center}

 \url{nbgrader.readthedocs.io}\qquad \url{github.com/jupyter/nbgrader/}
\end{frame}

\begin{frame}
 \begin{center}
  \begin{minipage}{0.6\textwidth}
   \tableofcontents
  \end{minipage}
 \end{center}
\end{frame}

\section{Jupyterhub and Jupyter notebooks}

\begin{frame}
 \begin{center}
  \begin{minipage}{0.6\textwidth}
   \tableofcontents[currentsection]
  \end{minipage}
 \end{center}
\end{frame}

\begin{frame}{Jupyterhub and nbgrader}

 \begin{center}
 \includegraphics[width=0.9\textwidth]{jupyterhub_nbgrader}
 \end{center}
 \begin{itemize}
  \item easily accessible interface to problem sets
  \item no need to install Python on local computer
  \item consistent working environment for all students
  \item \but may be beginners should gather experience with Python
	on their own computer
 \end{itemize}
\end{frame}

\begin{frame}{Jupyter notebooks or something else?}
 \begin{columns}
  \begin{column}{0.6\textwidth}
   \includegraphics[width=\textwidth]{notebook}
  \end{column}%
  \begin{column}{0.4\textwidth}
   \includegraphics[width=\textwidth]{spyder}

   \includegraphics[width=\textwidth]{vim}
  \end{column}
 \end{columns}

 \vspace{0.5truecm}
 \begin{itemize}
  \item notebook allows to guide students through a problem set
  \item students might think that programming in Python implies
	working with a Jupyter notebook
 \end{itemize}
\end{frame}

\section{grading}

\begin{frame}
 \begin{center}
  \begin{minipage}{0.6\textwidth}
   \tableofcontents[currentsection]
  \end{minipage}
 \end{center}
\end{frame}

\begin{frame}{Grading}
 \begin{center}
  \includegraphics[width=0.8\textwidth]{answer_tests}
 \end{center}
 \begin{itemize}
  \item autograded answer
  \item autograder tests
  \item manually graded answer
 \end{itemize}
\end{frame}

\begin{frame}{Functions}
 \begin{center}
  \includegraphics[width=0.7\textwidth]{answer_tests_function}
 \end{center}

 \vspace{-0.4truecm}
 \begin{itemize}
  \item need to introduce functions very early in the course\\
	\but special aspects (no arguments, no return value, default arguments,
	keyword arguments, \dots) not needed
  \item students get used to logically structured code early on\\
	\but they do not do it themselves
  \item include docstrings to make task well defined\\
	students get used to the idea that a function contains a docstring\\
        \but they are not writing the docstring by themselves
 \end{itemize}
\end{frame}

\begin{frame}{Grading with tests}
 \begin{itemize}
  \item test-driven development\\
	\but tests are not developed by the students
  \item tests allow to give feedback through error messages\\
	\but students tend to rely on this feedback instead of developing their
	own critical view on their code
  \item ``Stupid mistakes'' will be made not only at the beginning
	\begin{itemize}
         \item ``trivial'' standard tests (are results returned, do they have the correct type, \dots?)
	 \item + tests of specific functionality which should not disclose the solution
	\end{itemize}
 \end{itemize}
 \begin{center}
  \includegraphics[width=\textwidth]{answer_tests_unittest1}
 \end{center}
\end{frame}

\section{branching paths}

\begin{frame}
 \begin{center}
  \begin{minipage}{0.6\textwidth}
   \tableofcontents[currentsection]
  \end{minipage}
 \end{center}
\end{frame}

\begin{frame}{Coping with heterogeneity}
 \begin{itemize}
  \item students without programming experience and (sometimes) code writing apprehension
  \item students with knowledge in another programming language\\
	code often does not look very pythonic\\
	example: loop over objects vs. loop over indices
 \end{itemize}

@@

 \begin{itemize}
  \item \but scientific applications are often considered as an extra mental burden
 \end{itemize}
\end{frame}

\begin{frame}{Examples of problem sets}

@@

\begin{frame}{Branching paths}
 \only<1>{%
  \begin{columns}
   \begin{column}{0.3\textwidth}
    \includegraphics[height=0.8\textheight]{branching_1}
   \end{column}%
   \begin{column}{0.7\textwidth}
    \includegraphics[width=\textwidth]{JuliaSet_screenshot}
   \end{column}
  \end{columns}
 }%
 \only<2>{%
  \begin{columns}
   \begin{column}{0.3\textwidth}
    \includegraphics[height=0.8\textheight]{branching_2}
   \end{column}%
   \begin{column}{0.7\textwidth}
    \includegraphics[width=\textwidth]{ColorRepresentation_screenshot}
   \end{column}
  \end{columns}
 }%
 \only<3>{%
  \begin{columns}
   \begin{column}{0.3\textwidth}
    \includegraphics[height=0.8\textheight]{branching_3}
   \end{column}%
   \begin{column}{0.7\textwidth}
    \includegraphics[width=\textwidth]{Grid_screenshot}
   \end{column}
  \end{columns}
 }%
 \only<4>{%
  \begin{columns}
   \begin{column}{0.3\textwidth}
    \includegraphics[height=0.8\textheight]{branching_4}
   \end{column}%
   \begin{column}{0.7\textwidth}
    \includegraphics[width=\textwidth]{JuliaIteration_screenshot}
   \end{column}
  \end{columns}
 }%
 \only<5>{%
  \begin{columns}[b]
   \begin{column}{0.3\textwidth}
    \includegraphics[height=0.8\textheight]{branching_4}
   \end{column}%
   \begin{column}{0.7\textwidth}
    \begin{itemize}
     \item individual path through problem possible
     \item \but notebooks are opened in new tabs, it is easy to lose track
    \end{itemize}

    \vspace{1truecm}
   \end{column}
  \end{columns}
 }
\end{frame}
